### 2019-02-13 Reporte_Tecnico_1st_year/Technical_Report_1st_Year.tex
cambios recientes

Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting (\cite{sutton1998reinforcement}), but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}). A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure. A causal structure encondes a series of \textit{cause-effect} relations between events and knowing such relations allows a decision maker to add extra knowledge into the uncertainty of his environment and also allows to plan ahead his actions since he can predict what a certain action will cause (\cite{spirtes2000causation}, \cite{pearl2018theoretical}).\\
\\
Causal reasoning is found at the very core of human reasoning, since it allows us to manipulate our environment and being able to predict effects of a given action (\cite{spirtes2000causation}). Causal claims also support reasoning of the form \textit{if...hadn't occurred then...wouldn't have happened}. It is argued by \cite{pearl2018why} that imagining alternative scenarios is fundamental for humans to make causal reasoning.  Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect (\cite{rottman2014reasoning}), our hypothesis is that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.

@@

Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed. A causal model is able to \textit{explain} why a particular decision was made since it allows to extract the causes and effects of an agent's actions. Once a causal model is acquired, an external user is able to reason about \textit{what...if...} statements that associative methods can not answer. When a decision maker chooses an action out of many, a causal model allows to ask what would've happened if another action was taken without actually performing the alternative action. It is important to be careful when using the word \textit{explanation} since it allows for several uses in language, here we understand \textit{explanation} as used by \cite{woodward2005making} who considers explanations where what is explained depends on other factors via some relationship that holds as a matter of empirical fact rather than \textit{logical} reasons.

Learning a causal model of a certain environment would allow transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
\subsection{Outline}
In the Section \ref{Main_Concepts} the main concepts that serve as a foundation for this work will be described; in particular, we will describe in detail what causality means. Since this research proposal asumes that a causal model governs an environment, a precise definition of causality will be given along with the mathematical machinery to perform causal inference. Also, since the entity that will be facing causal environments is a \textit{rational} decision maker we will precisely define, axiomatically, what does rationality means and which are the main theories of rational decision making. At the end of the section, we will describe a mathematical theory which is suitable to modelling interaction between rational decision makers since interactive learning is considered.

@@

\newpage
\section{Main Concepts}{\label{Main_Concepts}}
	\subsection{Attempts to define Causality}
	Causality has been a difficult concept to define and several attempts have been made. In fact, in mainstream statistics the term \textit{cause} have been nearly banned as documented by \cite{pearl2018why}.

@@

Since causal claims in common language can have different logical form, in this work we give a \textit{working} definition of causality that allows a formal language for stating and manipulating causal concepts and we will take the given definition as axiomatic. 

@@

	It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph (\cite{sucar2015probabilistic}).
	\subsubsection{The identifiability problem}
	Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data? It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) (\cite{peters2012identifiability}, \cite{mooij2016distinguishing}).
	\subsubsection{Do-Calculus}
The Do-Calculus (\cite{pearl1995causal}, \cite{pearl2009causality}) is a set of rules for manipulating probabilistic statements that involve interventions and, under certain conditions, allow them to be transformed into statements that do not involve interventional data.

@@

\subsubsection{Use of Causal Graphical Models over other causal models}
We choose to use specifically a Causal Graphical model because of the mathematical machinery that exists around them and because Graphs are a model of Cognitive Processes, such as causal reasoning (\cite{glymour2003learning}, \cite{danks2014unifying}). In this way, if an agent \textit{knows} that a certain environment is governed by causal relations, then a causal graphical model can be used to model the causal reasoning of the agent.

@@

	On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain. It allows to ask \textit{why}.

@@

	\subsection{Research Question}
The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
\begin{enumerate}
\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices? \item How can a causal structure help a decision maker in order to guide his learning process? 
\item What does the rationality assumption implies about how to choose when considering causal information? 
\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
\end{enumerate}
	\subsection{Hypothesis}
	Let $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn a causal model of the environment and use it to find the optimal action (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
	\subsection{Objectives}
		\subsubsection{General Objectives}
			The proposed research has as a general objective to understand what are the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
		\subsubsection{Specific Objectives}
		\begin{enumerate}
		\item In order to achieve the general objective, we need to separate it into smaller problems which are required to be solved first. In first place, it is required to find a general way of defining a game that captures the interaction between a decision maker and a causal-governed environment.

		\item In second place, to find a family of distributions that represent any past knowledge and current beliefs about the causal structure of the environment and a way to use these beliefs in order to make choices according to the decision maker's preferences.

		\item Beliefs must be updated after external information is provided by the environment in response to any given action that was taken by the decision maker. A general updating criteria must be provided which uses the causal nature of the environment.
		\end{enumerate}

@@

		\subsubsection{Extensions}
		Given the Limitations addressed in the previous section, some possible extensions could be:
		\begin{itemize}
		\item Consider longer games.
		\item Consider more players than an agent and Nature.
		\item Consider objectives/intentions for Nature.
		\item Studying game-theoretical properties of the model such as equilibria in terms of the causal structure.
		\end{itemize}
\newpage
\section{Related Work}{\label{Related_Work}}
Reinforcement Learning (RL) has been the standard setting for decision-making and learning by interaction. In this setting it is assumed that a Markov Decision Process (MDP) is able to capture the effects of the agent's actions in the environment. An MDP is composed by a set of states, a transition function and a reward function, where an agent's action moves the state of the environment according to the transition function and obtaining a reward in such process.

@@

In this research we are looking for a methodology that will allow to discover causal structure from an uncertain environment while also using it to obtain a reward. Being able to discover the causal structure at the same time that it is being used will allow to eliminate the restriction of knowing the full causal model or part of it. 
\newpage
\section{Methodology}
In this section we describe the steps that are to be followed in order to answer the Hypothesis.
\subsection{A guiding principle for a Causal Decision Problem}
We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational. We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
\begin{prop}
Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent. Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
\end{prop}
\subsection{Modelling a Decision Problem under Uncertainty as a Game}
Consider a decision problem under uncertainty $(\mathcal{A},\mathcal{E},\mathcal{C},\succeq)$ where the available actions $\mathcal{A}$ to a decision maker are causally related to the outcomes $\mathcal{C}$ and this causal relations are captured by a Causal Graphical Model $\mathcal{G}$. The Causal model is assumed to remain fixed and unknown to the decision maker, who is assumed to be rational and aware of the causal nature of the decision problem he faces.

@@

The chosen actions will be compared to classic on-line decision making 
\subsubsection{Internal Validation: Reproducibility}
For the results obtained by following the methodology described in previous sections be reproducible, the causal model used to simulate an environment must be fully specified by describing the probabilistic parameters that control it. The initial distributions given to the decision maker to encode his beliefs must be shown as well with the updating criteria used when information from the environment becomes available. 
\subsubsection{External Validation: Different problems}
For the proposed method to be tested in different decision problems, it is required that the correspondent causal model satisfies all of the conditions previously described, and any derived algorithm will be designed in such a way that the characteristics of the causal model are an external input. In this way, if an external user is able to give a causal model, the method is expected to work independent of the particularities of the causal model which is being tested.
\subsubsection{Concurrent Validation}
Concurrent validation will be tested by comparing the performance obtained by the proposed method against the theoretical optima.
\subsection{Working plan}
Concrete steps towards solving the problem:
\begin{itemize}
\item Literature review, understanding the problem, considering several models and possible solutions.
\item Studying implications of the rationality assumption for decision making under uncertainty in causal environments.
\item Formalization of a Causal Decision Problem: a decision problem under uncertainty where causal relations are encoded into a Causal Graphical Model. Propose a solution using the rationality assumption.
\item Propose a representation of beliefs that allow causal structure to be considered. Three subcases considered
\item Propose and evaluate an updating criteria for causal beliefs that can be updated according to new information.
\item Study the problem of on-line causal discovery.
\item Considering interventions in more than one variable.

\end{itemize}
\subsubsection{Publications plan}
\begin{itemize}
\item \textbf{Year 1:} Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}. Accepted for poster presentation at the \textbf{CausalML 2018 Workshop at ICML 2018}. July 15 2018, Stockholm, Sweden. This can be confirmed at \url{https://sites.google.com/site/faim18wscausalml/program-schedule}. The paper can be found at \url{https://arxiv.org/abs/1807.01268}.
\item \textbf{Year 2:} Conference Paper
		\begin{itemize}
		\item The implications of Causal Relations for Decision Making: a general principle for Decision under Uncertainty.
		\item Possible Forum: Conference on Uncertainty in Artificial Intelligence (UAI), International Conference Machine Learning (ICML)
		\end{itemize}
\item \textbf{Year 2:} Conference Paper
	\begin{itemize}
	\item When a causal model is unknown, how to encode uncertainties about causal structure in a probabilistic distribution? Bayesian Nonparametric prior that explicitly represents causal structure.
	\item Possible Forum: International Conference on Bayesian Nonparametrics, International Conference on Probabilistic Graphical Models.
	\end{itemize}
\item \textbf{Year 3:} Journal Paper
		\begin{itemize}
			\item Aim: Theoretical paper of results for on-line causal learning.
		\end{itemize}
\end{itemize}

@@

@@

Belief updating using causal information when the decision maker doesn't know the structure of the model nor its parameters is yet to be studied and left as future work.

%------------- References --------------------
\singlespacing
%doi=false,isbn=false,url=false
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}

>>>>

Learning to perform tasks by interaction has been extensively studied in Computer Science using Reinforcement Learning (RL) (\cite{sutton1998reinforcement}), but the most common used techniques in this field are based on purely associative relations and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}). A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure. A causal structure encondes a series of \textit{cause-effect} relations between events and knowing such relations allows a decision maker to add extra knowledge into the uncertainty of his environment and also allows to plan ahead his actions since he can predict what a certain action will cause (\cite{spirtes2000causation}, \cite{pearl2018theoretical}).\\
\\
Causal reasoning is found at the very core of human reasoning (\cite{danks2014unifying}, \cite{lake2017building}), since it allows us to manipulate our environment and being able to predict effects of a given action (\cite{spirtes2000causation}). It is even argued that in a more basic level the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} (\cite{friston2010free}, \cite{hohwy2013predictive}, \cite{clark2015surfing}). Causal claims also support reasoning of the form \textit{if...hadn't occurred then...wouldn't have happened}. It is argued by \cite{pearl2018why} that imagining alternative scenarios is fundamental for humans to make causal reasoning.  Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect (\cite{rottman2014reasoning}), our hypothesis is that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.

@@

Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed. A causal model is able to \textit{explain} why a particular decision was made since it allows to extract the causes and effects of an agent's actions. Once a causal model is acquired, an external user is able to reason about \textit{what...if...} statements that associative methods can not answer. When a decision maker chooses an action out of many, a causal model allows to ask what would've happened if another action was taken without actually performing the alternative action. It is important to be careful when using the word \textit{explanation} since it allows for several uses in language, here we understand \textit{explanation} as used by \cite{woodward2005making} who considers explanations where what is explained depends on other factors via some relationship that holds as a matter of empirical fact rather than \textit{logical} reasons. Also, learning a causal model of a certain environment would allow transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
\subsection{Outline}
In the Section \ref{Main_Concepts} the main concepts that serve as a foundation for this work will be described; in particular, we will describe in detail what causality means. Since this research proposal asumes that a causal model governs an environment, a precise definition of causality will be given along with the mathematical machinery to perform causal inference. Also, since the entity that will be facing causal environments is a \textit{rational} decision maker we will precisely define, axiomatically, what does rationality means and which are the main theories of rational decision making. At the end of the section, we will describe a mathematical theory which is suitable to modelling interaction between rational decision makers since interactive learning is considered.

@@

\newpage
\section{Main Concepts}{\label{Main_Concepts}}

\subsection{Causality}
Since Causality is at the core of this proposal, we must make a careful treatment of it, both in its definition and use. Leaving aside deeper philosophical and epistemological issues, we consider \textit{causal structre} as part of of the world itself, something that holds deep inside the tissue of the physical world. We consider that the notion of causal relations does not contradict the  stochastic nature of the world in the sense of \citep{gilboa2009decision}, since we can both know that throwing a dice is both a stochastic phenomenon while also acknowledging that throwing it caused a stochastic quantity to be realized.

\subsection{Difference between Causal Relations and Associative Relations}
As \cite{pearl2018why} mentions, there are three levels of causal reasoning: observing, manipulating and counterfactual reasoning. Associative relations belong to the first level, while causal reasoning encompasses the second and third levels the core of reasoning with causal relations. Associative relations are based on correlation between events, or patterns of occurence to be found in data, while causal relations are based in cause-efect patterns, which result from some mechanism which is rooted in the nature of the observed events. As an example, although a very simple one, consider the following: a dog that lives as a pet in some household has observed that every day he has food so he does not worry about eating or not, but any kid knows that whether he eats or not depends on certain grown-up being present. The dog uses only observed occurrences in order to obtain conclusions, while the kid can imagine what would happen if his parents arrive home late.

It is important to mention that although in the usual examples of causal relations the cause precedes temporally the effect this need not be a general rule. It suffices to show one example of a causal relation in which the effect precedes temporally the cause in order to conclude that cause-effect relations need not be in temporal order. For example, consider the example of a barometer and a storm; in this case, the barometer drops first, but it is the storm which actually causes this event (\cite{pearl2009causality}).

\subsection{About definitions of Causality}
Causality has been a difficult concept to define and several attempts have been made. In fact, in mainstream statistics the term \textit{cause} have been nearly banned as documented by \cite{pearl2018why}.

@@

Since causal claims in common language can have different logical form, in this work we give a \textit{working} definition of causality that allows a formal language for stating and manipulating causal concepts and we will take the given definition as axiomatic. From this point, the words \textit{causality}, \textit{causal relation} and similar terms will be understood in terms of the following formal definition and not to be confused with the ordinary-language words that are spelled in the same way. 

@@

It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph (\cite{sucar2015probabilistic}).
\subsubsection{The identifiability problem}
The identifiability problem is to find under what are the conditions under which causal inquiries can be answered in terms of purely \textit{observational} data? It has been shown (\cite{peters2012identifiability}, \cite{mooij2016distinguishing}) that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence).
\subsubsection{Do-Calculus}
The Do-Calculus (\cite{pearl1995causal}, \cite{pearl2009causality}) is a set of rules for manipulating probabilistic statements that involve interventions and, under certain conditions, allow them to be transformed into statements that do not involve interventional data.

@@

\subsubsection{Use of Causal Graphical Models over other causal models}
We choose to use specifically a Causal Graphical model because of the mathematical machinery that exists around them and because Graphs are a model of Cognitive Processes, such as causal reasoning (\cite{glymour2003learning}, \cite{danks2014unifying}). In this way, if an agent \textit{knows} that a certain environment is governed by causal relations, then a causal graphical model can be used to model the process of causal reasoning of the agent in a way that resembles human reasoning.

\subsection{Advantages of Causal Models}
If a Causal Model is learned from data, and if it is a \textit{good} model, then according to Pearl's levels, the model can be used to predict the effect of certain interventions as well to perform counterfactual reasoning in order to evaluate an intervention that has been made; i.e., what if another intervention had been performed?

A Causal Model can allow for transferability of knowledge to similar domains.

@@

	On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain. It allows to ask \textit{why}. Since a causal model, once it is learned, does not depend neither on the agent nor its preferences, the gained knowledge could be easily transfered for problems of similar domain where the use of the acquired causal relations is useful.

@@

\subsection{Research Question}
The proposed research attempts to answer a series of questions about the nature of causal relations and about their possible use.
\begin{enumerate}
\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices? \item How can a causal structure help a decision maker in order to guide his learning process? 
\item When does causal information will be helpful?
\item What does the rationality assumption implies about how to choose when considering causal information? 
\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
\end{enumerate}
\subsection{Hypothesis}
Let $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn a causal model of the environment and use it to find the optimal action (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
\subsection{Objectives}
\subsubsection{General Objectives}
The proposed research has as a general objective to provide the understanding and knowledge about what are the implications of causality for rational decision maker who faces an uncertain environment and about how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
\subsubsection{Specific Objectives}
In order to achieve the general objective, we need to separate it into smaller problems which are required to be solved first. Given a scenario which can be modelled as a causal decision problem:
\begin{enumerate}
\item In first place, we require to characterize the problems where this approach can be used.

\item  In second place, it is required to find a general way of defining a game that captures the interaction between a decision maker and a causal-governed environment.

\item In third place, to find a family of distributions that represent any past knowledge and current beliefs about the causal structure of the environment and a way to use these beliefs in order to make choices according to the decision maker's preferences.

\item Beliefs must be updated after external information is provided by the environment in response to any given action that was taken by the decision maker. A general updating criteria must be provided which explicitly uses the causal nature of the environment.
\end{enumerate}

@@

\subsubsection{Domain of application}
Even if, as argued in the Introduction, causal relations are present everywhere, we do not pretend that our proposed method can be applied anywhere without further considerations. For the proposed method to be applied, it is necessary that a Decision Problem exists (i.e., a non-empty set of available actions exist which is related to a well-defined goal to be achieved), and that the elements of the decision problem can be causally related through a causal model. This is, we require:
\begin{enumerate}
\item The main task is to choose an action in order to achieve some objective.
\item The set of available actions must have at least one element which is known by the decision maker. 
\item Rational preferences for the decision maker. 
\item Known environment.
\item Causal relations in the environment. 
\item Decision-making steps in discrete time. 
\end{enumerate}

\subsubsection{Extensions}
Given the Limitations addressed in the previous section, some possible extensions could be:
\begin{itemize}
\item Consider longer games.
\item Consider more players than an agent and Nature.
\item Consider objectives/intentions for Nature.
\item Studying game-theoretical properties of the model such as equilibria in terms of the causal structure.
\end{itemize}

\newpage

\section{Related Work}{\label{Related_Work}}
Reinforcement Learning (RL) has been the standard setting for decision-making and learning by interaction. In this setting it is assumed that a Markov Decision Process (MDP) is able to capture the effects of the agent's actions in the environment. An MDP is composed by a set of states, a transition function and a reward function, where an agent's action moves the state of the environment according to the transition function and obtaining a reward in such process.

@@

In this research we are looking for a methodology that will allow to discover causal structure from an uncertain environment while also using it to obtain a reward. Being able to discover the causal structure at the same time that it is being used will allow to eliminate the restriction of knowing the full causal model or part of it. 
\newpage
\section{Methodology}
In this section we describe the steps that are to be followed in order to answer the Hypothesis.
\subsection{A guiding principle for a Causal Decision Problem}
We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational. We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
\begin{prop}
Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $Y=1$ is the most desired outcome for the agent. Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}, \]
where $P$ is a probability distribution which encodes the agent knowledge, or beliefs, about the causal relations that hold in his evironment. This is equivalent to finding $a^\ast$  as
\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g). \]
Since the action chosen is the action with the highest probability of producing the most desired action, then it is the action that maximizes the expected utility for the agent.\\
\\
While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
\[ U(x)=\sum_y P(y | do(x)) u(y), \]
we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment. We note that our proposal makes use of an implicit probability distribution over causal structures, since the decision maker holds beliefs about the causal nature of his environment and within each realization of those beliefs the maximization for the optimal action is performed, while Pearl's function assumes a single causal structure.
\end{prop}

\subsection{Modelling a Decision Problem under Uncertainty as a Game}
Consider a decision problem under uncertainty $(\mathcal{A},\mathcal{E},\mathcal{C},\succeq)$ where the available actions $\mathcal{A}$ to a decision maker are causally related to the outcomes $\mathcal{C}$ and this causal relations are captured by a Causal Graphical Model $\mathcal{G}$. The Causal model is assumed to remain fixed and unknown to the decision maker, who is assumed to be rational and aware of the causal nature of the decision problem he faces.

@@

The chosen actions will be compared to classic on-line decision making 
\subsubsection{Internal Validation: Reproducibility}
For the results obtained by following the methodology described in previous sections be reproducible, the causal model used to simulate an environment must be fully specified by describing the probabilistic parameters that control it. The initial distributions given to the decision maker to encode his beliefs must be shown as well with the updating criteria used when information from the environment becomes available. 
\subsubsection{External Validation: Different problems}
For the proposed method to be tested in different decision problems, it is required that the correspondent causal model satisfies all of the conditions previously described, and any derived algorithm will be designed in such a way that the characteristics of the causal model are an external input. In this way, if an external user is able to give a causal model, the method is expected to work independent of the particularities of the causal model which is being tested.
\subsubsection{Concurrent Validation}
Concurrent validation will be tested by comparing the performance obtained by the proposed method against the theoretical optima.
\subsection{Working plan}
Concrete steps towards solving the problem:
\begin{itemize}
\item Literature review, understanding the problem, considering several models and possible solutions.
\item Studying implications of the rationality assumption for decision making under uncertainty in causal environments.
\item Formalization of a Causal Decision Problem: a decision problem under uncertainty where causal relations are encoded into a Causal Graphical Model. Propose a solution using the rationality assumption. 
\item Propose a representation of beliefs that allow causal structure to be considered. Three subcases considered. Conference Paper No. 1
\item Propose and evaluate an updating criteria for causal beliefs that can be updated according to new information when involved variables are known. Conference Paper No. 2
\item Study the problem of on-line causal discovery (fully unknown model). Convergence of proposed solution. Conference Paper No. 3 and Journal Paper.
\item Considering interventions in more than one variable.

\end{itemize}
\subsubsection{Publications plan}
\begin{itemize}
\item \textbf{Year 1:} Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}. Accepted for poster presentation at the \textbf{CausalML 2018 Workshop at ICML 2018}. July 15 2018, Stockholm, Sweden. This can be confirmed at \url{https://sites.google.com/site/faim18wscausalml/program-schedule}. The paper can be found at \url{https://arxiv.org/abs/1807.01268}.
\item \textbf{Year 2:} Conference Paper
		\begin{itemize}
		\item The implications of Causal Relations for Decision Making: a general principle for Decision under Uncertainty.
		\item Possible Forum: Conference on Uncertainty in Artificial Intelligence (UAI), International Conference Machine Learning (ICML)
		\end{itemize}
\item \textbf{Year 2:} Conference Paper 2 and 3
	\begin{itemize}
	\item When a causal model is unknown but the involved variables are known, how to encode uncertainties about causal structure in a probabilistic distribution? Bayesian Nonparametric prior that explicitly represents causal structure.
	\item When variables of the environment are unknown how to encode prior beliefs and how to update them when data is not fully observed.
	\item Possible Forum: International Conference on Bayesian Nonparametrics, International Conference on Probabilistic Graphical Models.
	\end{itemize}
\item \textbf{Year 3:} Journal Paper
		\begin{itemize}
			\item Aim: Theoretical paper of results for on-line causal learning.
		\end{itemize}
\end{itemize}

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[scale=0.5]{Gantt.eps}
\caption{Gantt Diagram for the proposed research}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}

@@

We observe that when the number of rounds is small, as in Figure 2, Q-learning outperforms our method, which achieves similar performance than the random-acting agent. On the other hand, in 50 rounds (Figure 3), our Causal Agent and the Q-Agent have left behind the Random Agent, which is even more noticeable by 200 rounds. This means that our agent (Causal Agent) is both learning a good action while learning a Causal Model of the environment that he is facing.

@@

Belief updating using causal information when the decision maker current knowledge is specified as a Bayesian Nonparametric Prior (BNP)  is yet to be studied and left as future work, since the construction and use of this kind of priors is a hard problem, this is because a BNP is basically a probability distribution over infinite-dimensional spaces, such as the space of continuous functions or the space of unbounded graphs (\cite{hjort2010bayesian}, \cite{phadia2015prior}, \cite{ghosal2017fundamentals}).
\newpage
\section{Conclusions}
We have proposed that an autonomous, rational agent who faces an uncertain environment which is governed by an unknown causal mechanism can make use of the causal relations that hold in such environment in order to find an action that will take him into the way of a desired result, in terms of his (rational) preferences. Also, causal information can be acquired while interacting with the environment and this newly acquired information will enrich the decision making procedure.

We have taken inspiration in how human beings use causal information when interacting with the world, which has a complex causal structure, in order to pursue objectives. Following this intuitions we proposed a \textit{guiding principle} to exploit causal information that must be followed by a rational agent in order to be coherent with the definition of rationality. From the guiding principle we stated a systematic approach into causal decision making and considered three cases to be solved in order. In two of the three cases we showed that a performance similar to classical decision-making algorithms can be achieved while also learning a causal model from the environment. We are considering. If this research is allowed to continue, we will delve into general settings that resemble realistic situations in which the causal model that controls an environment is barely known.
\newpage
%------------- References --------------------
\singlespacing
%doi=false,isbn=false,url=false
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
